---
title: "Statia Snake"
author: "D. S. Fernandez del Viso"
date: "2/28/2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading data
Data from Hanna.  
_Note: occupancy = occurrence_. 

> Questions: Each year is a season, with many surveys?  Or, each survey is a season, or each month is a season, with several surveys?  
> 2018 and 2019 with the same number of surveys?  


```{r}
library(readxl)
Also2018 <- read_excel("Alsophis 2018 occupancy data.xlsx")
head(Also2018)
Also2019 <- read_excel("Alsophis 2019 occupancy data.xlsx")
head(Also2019)
```

## Cleaning the names of the columns  
Note: _Why to "clean" variable names? The space?_  

```{r}
# install.packages("janitor")
library(janitor)
Also2018 <- clean_names(Also2018)
head(Also2018)
Also2019 <- clean_names(Also2019)
head(Also2019)
```

## Using occSS
__occSS__: allows for psi or p to be modelled as a logistic function of site covariates or survey covariates, as specified by model. It includes a built in .time covariate which can be used for modelling p with time as a fixed effect, and .Time for a linear or quadratic trend. A built-in .b covariate corresponds to a behavioural effect, where detection depends on whether the species was detected on the previous occasion or not.  

_link_: function to link the response on a (0,1) scale to a linear predictor on (-Inf, Inf). The canonical link is the logistic ("logit") function, which has some nice theoretical properties and can be interpreted as the log of the odds of success. Other links are available, notably the cumulative standard normal ("probit") link, which allows for Gibbs sampling with truncated normal distributions. For that reason, several of the Bayesian estimation functions in wiqid use the probit link.  

Note: _Why Also24?_

```{r}
library(wiqid)
#with link 
#2018 logit
occSS(Also2018, link=c("logit"))
#2019 logit
occSS(Also2019[,1:17], link=c("logit"))
#2018 probit
occSS(Also2018, link=c("probit"))
#2019 probit
occSS(Also2019[,1:17], link=c("probit"))
#A built-in .b covariate corresponds to a behavioural effect, where detection depends on whether the species was detected on the previous occasion or not.
#2018 
occSS(Also2018, p ~ .b)
#2019
occSS(Also2019[,1:17], p ~ .b)
```

_occSS0_:implements a simple model with one parameter for probability of occupancy and one for probability of detection, ie. a psi(.) p(.) model.

```{r}
#2018
# with Also24 1:17
Also24 <- Also2018[,1:17]
y <- rowSums(Also24, na.rm=TRUE)
n <- rowSums(!is.na(Also24))
# occSS0 is a faster simple form of occSS
occSS0(y, n)
#2019
y <- rowSums(Also2019[,1:17], na.rm=TRUE)
n <- rowSums(!is.na(Also2019[,1:17]))
# occSS0 is a faster simple form of occSS
occSS0(y, n)
```

occSS and occSS0 do the same thing for 2019
but for 2018 depend if we use all data or 1:17

##Other occSSs  

__occSStime__: allows for time-varying covariates that are the same across all sites, eg, moon-phase. A categorical time variable .time and a time trend .Time are built-in. A plot of detection probability vs time is produced if plot=TRUE. 

Note: _Does not work with Also2018_

```{r}
#time effect accross all sites
#2018
time <- occSStime(Also24, p ~ .time)
time[["real"]]
# trend linear
Time <- occSStime(Also24, p ~ .Time)
Time[["real"]]
# trend quadratic
Time2 <- occSStime(Also24, p ~ .Time + I(.Time^2))
Time2[["real"]]
#2019
time <- occSStime(Also2019[,1:17], p ~ .time)
time[["real"]]
# trend linear
Time <- occSStime(Also2019[,1:17], p ~ .Time)
Time[["real"]]
# trend quadratic
Time2 <- occSStime(Also2019[,1:17], p ~ .Time + I(.Time^2))
Time2[["real"]]
```

Apparently it is nor worthwhile to do more than 3 - 5 surveys on a year/season according to 2018, but maybe is good to do 10 surveys according to 2019.

__occSScovSite__: allows for site covariates but not for occasion or survey covariates.

```{r}
Also2019_17 <- Also2019[,1:17]
y <- rowSums(Also2019_17, na.rm=TRUE)
n <- rowSums(!is.na(Also2019_17))
# model habitat
occSScovSite(y, n, model = psi ~ habitat, data = Also2019)
# model elevation
occSScovSite(y, n, model = psi ~ elevation, data = Also2019)
res2 <- occSScovSite(y, n, model = psi ~ elevation, data = Also2019)
# plot psi vs elevation
psi <- as.data.frame(res2[["real"]])
psi <- psi[1:107,1]
plot(Also2019$elevation,psi)
```

Apparently low elevation sites tend to have more probability of being occupied.

Need file for 2018, or is the same structure? (same sequence of sites)

Note: _Do we have multi season-month or multi years?_

##Royle-Nichols functions  

These functions implement the Royle-Nichols method (Royle & Nichols 2003) for estimation of site occupancy allowing for abundance-induced heterogeneity in detection probability. Probability of detection is modelled as a function of the number of animals available for detection, n, and the probability of detection of an individual animal, r. Probability of occupancy is derived as the probability that n > 0.

Function occSSrn allows for site-specific covariates to be included in the model. occSSrnSite and occSSrn0 are fast alternatives that do not require a full detection history matrix.
```{r}
# 2018
#basic model
Also24 <- Also2018[,1:17]
occSSrn(Also24)
# 2019
#basic model
Also2019_17 <- Also2019[,1:17]
occSSrn(Also2019_17)
#lambda (and psi?) - elevation
Also2019_17 <- Also2019[,1:17]
occSSrn(Also2019_17, model = lambda ~ elevation, data = Also2019)
#lambda (and psi?) - habitat
Also2019_17 <- Also2019[,1:17]
occSSrn(Also2019_17, model = lambda ~ habitat, data = Also2019)
#r - elevation
Also2019_17 <- Also2019[,1:17]
occSSrn(Also2019_17, model = r ~ elevation, data = Also2019)
```

##Graphs

```{r}
Also2019_17 <- Also2019[,1:17]
lamelev <- occSSrn(Also2019_17, model = lambda ~ elevation, data = Also2019)
lamelev <- as.data.frame(lamelev[["real"]])
#psi vs elevation
plot(Also2019$elevation,lamelev[1:107,1])
#lambda vs elevation
plot(Also2019$elevation,lamelev[108:214,1])
###
Also2019_17 <- Also2019[,1:17]
relev <- occSSrn(Also2019_17, model = r ~ elevation, data = Also2019)
relev <- as.data.frame(relev[["real"]])
#r vs elevation
plot(Also2019$elevation,relev[215:321,1])
```





++++++++++++++++++++++++++++
EXAMPLE MS
============================



##GrandSkins data set

```{r}
data(GrandSkinks)
head(GrandSkinks)

#testing
data(GrandSkinks)
DH <- GrandSkinks[, 1:15]

occMS0(DH, 3)

occMStime(DH, 3, model=list(gamma ~ .interval, epsilon~1, p~.season))

occMScovSite(DH, 3,
   model=list(psi1~habitat, gamma ~ .interval, epsilon~habitat, p~.season),
   data=GrandSkinks)

occMS(DH, 3,
      model=list(psi1~habitat, gamma ~ .interval, epsilon~habitat, p~.season),
   data=GrandSkinks)
```

